\section{Algorithmic Backends}\label{sec:al}
This section presents a mechanism for automatically generating
algorithmic backends from the formal semantics.
We define \emph{\dl}, a declarative language that defines the formal
semantics of Wasm (Sec.~\ref{sec:dl}),
\emph{\al}, an algorithmic language that defines the Wasm semantics
in a pseudocode style (Sec.~\ref{sec:aldef}),
and a \dl to \al translation (Sec.~\ref{sec:dl2al}).
We then describe generating a prose specification from the semantics
described in \al (Sec.~\ref{sec:prose}) and interpreting the prose
specification which enables the indirect interpretation of Wasm (Sec.~\ref{sec:interp}).

\subsection{\dl: Declarative Language}\label{sec:dl}
The DSL describes the formal semantics of Wasm, and
we abstract it into a \dl that only shows the features relevant to translation to an algorithmic representation.
Fig.~\ref{fig:dl-syntax} presents the syntax of \dl.

\begin{figure}[t]
\[
\small
\begin{array}{l@{~}c@{~}r@{~}l}
\text{Semantics} & \delta^*\\
\text{Definition} & \delta &::=& \rho\ \mid\ \lambda\\
\text{Reduction rule} & \rho &::=& \gamma \leadsto \gamma\ \mbox{---}\ \pi^*\\
\text{Configuration} & \gamma &::=&(\eta_\bot,\ \eta)\\
\text{Premise} & \pi &::=& \eta\ \mid\ \mathsf{\small otherwise}\\
\text{Helper function} &
\lambda &::=& \varx \kwrl \eta^* \kwrr\ \kwequ\ \eta\ \mbox{---}\ \pi^*\\
\text{Expression} & \eta &::=&
x\ \mid\ n\
\mid\ \kappa\ \eta^*\ \mid\ \eta + \eta\
\mid \eta^\eta\ \mid (\eta,\ \eta)\ \mid \eta^*
\mid \eta\ = \eta\ \mid \eta \leftarrow \eta\
\mid \cdots \\
\text{Constructor} & \kappa &::=&
\ssf{I32} \mid\ \ssf{CONST}\ \mid\ \ssf{REF.IS\_NULL}\ \mid\ \ssf{REF.NULL}\
\mid\ \ssf{LABEL\_}\
\mid\ \ssf{BR}\
\mid \cdots
\end{array}
\]
\caption{Syntax of \dl for Wasm}\label{fig:dl-syntax}
\end{figure}

The Wasm semantics is defined by a sequence of definitions $\delta^*$.
A definition is either a reduction rule $\rho$ or an auxiliary helper function $\lambda$.
A reduction rule $\gamma_1 \leadsto \gamma_2\ \mbox{---}\ \pi^*$ denotes that
when the current configuration of a program matches $\gamma_1$ and
all $\pi^*$ are evaluated to true, then the program configuration becomes $\gamma_2$.
A configuration $(\eta_\bot,\ \eta)$ denotes an optional Wasm program state $\eta_\bot$,
which is a pair of the current store and the current frame,
and a list of Wasm instruction $\eta$ that represents the current stack.
A premise $\pi$ is a condition expression $\eta$ or a special keyword
\ensuremath{\mathsf{\small otherwise}},
which denotes the negation of all the previous premises.
A helper function $\varx \kwrl \eta^* \kwrr\ \kwequ\ \eta\ \mbox{---}\ \pi^*$ denotes that
when a function named $\varx$ is called, its arguments are bound to parameters $\eta^*$
and a body expression $\eta$ is evaluated if all $\pi^*$ are evaluated to true.
An expression $\eta$ is a DSL expression.
Because the details of the expression is not relevant to this section,
we show only some cases used for concrete
examples.
A constructor $\kappa$ is an atomic string that denotes the names of
Wasm types (i.e. $\ssf{I32}$) or Wasm instructions (i.e. $\ssf{REF.IS\_NULL}$).

For example, the semantics of \inblue{\ensuremath{\mathsf{ref.is\_null}}}
in Fig.~\ref{fig:dsl} corresponds to the following in \dl:
\[
\begin{array}{l@{}l@{~}c@{~}l}
[&(\bot, [\upsilon, \ssf{REF.IS\_NULL}]) &\leadsto& (\bot, [\ssf{CONST\ I32\ 1}])\
\mbox{---}\ [\upsilon = \ssf{REF.NULL}\ t],\\
&(\bot, [\upsilon, \ssf{REF.IS\_NULL}])&\leadsto&(\bot, [\ssf{CONST\ I32\ 0}])\
\mbox{---}\ [\ssf{\small otherwise}] ]
\end{array}
\]
where a sequence is represented as comma separated elements
enclosed by $[$ and $]$ for readability.

\subsection{\al: Algorithmic Language}\label{sec:aldef}
\al is an algorithmic language that defines the Wasm language semantics in a pseudocode style.
Definitions in \dl are translated into definitions in \al,
so that algorithmic backends can be generated.

\begin{figure}[t]
\[
\small
\begin{array}{l@{~}c@{~}r@{~}lll}
\text{Program} & \prog &::=& \alg^*\\
\text{Algorithm} & \alg &::=&
\kwalg \; \varx \kwrl \expr^* \kwrr \; \kwcl \inst^* \kwcr \\
\text{Instruction} & \inst &::=&
    \kwif \; \cond \; \inst^* \; \inst^* & \mbox{If $\cond$, then: $\inst_1^*$ Else: $\inst_2^*$}\\
&& \mid&
  \kweither \; \inst^* \; \inst^* & \mbox{Either: $\inst_1^*$ Or: $\inst_2^*$}\\
&& \mid&
  \kwenter \; \expr \; \expr \; \inst^* & \mbox{Enter $\expr_1$ with label $\expr_2$\ :\ $\inst^*$}\\
&& \mid&
  \kwassert \; \cond  & \mbox{Assert: Due to validation, $\cond$.}\\
&& \mid&
  \kwpush \; \expr  & \mbox{Push $\expr$ to the stack.}\\
&& \mid&
  \kwpop \; \expr  & \mbox{Pop $\expr$ from the stack.}\\
&& \mid&
  \kwpopall \; \expr  & \mbox{Pop all values $\expr$ from the stack.}\\
&& \mid&
  \kwlet \; \expr \; \expr & \mbox{Let $\expr_1$ be $\expr_2$.}\\
&& \mid&
  \kwtrap & \mbox{Trap.}\\
&& \mid&
  \kwnop & \mbox{Do nothing.}\\
&& \mid&
  \kwreturn \; \expr^? & \mbox{Return $\expr^?$.}\\
&& \mid&
  \kwexecute \; \expr & \mbox{Execute $\expr$.}\\
&& \mid&
  \kwexecuteseq \; \expr & \mbox{Execute the sequence $\expr$.}\\
&& \mid&
  \kwperform \; \varx \; \expr^* & \mbox{Perform $\varx(\expr^*)$.}\\
&& \mid&
  \kwexit & \mbox{Exit current context.}\\
&& \mid&
    \kwreplace \; \expr \; \qual \; \expr & \mbox{Replace $\expr_1[\qual]$ with $\expr_2$.}
\\

\text{Expression} & \expr &::=&
    \varx & \varx\\
&& \mid&
  \num & \num\\
&& \mid&
  \bcode{-} \; \expr & \bcode{-} \; \expr\\
&& \mid&
  \expr \; \binop \; \expr & \expr_1 \; \binop \; \expr_2\\
&& \mid&
  \expr \kwsl \qual \kwsr & \expr \kwsl \qual \kwsr\\
&& \mid&
  \expr \kwsl \qual \kwsr \; \kwass \; \expr & \mbox{$\expr_1$ with $\qual$ replaced by $\expr_2$}\\
&& \mid&
  \expr \kwsl \qual \kwsr \; \kwext \; \expr & \mbox{$\expr_1$ with $\qual$ prepended/appended by $\expr_2$}\\
&& \mid&
  \kwcl (\varx \mapsto \expr)^* \kwcr & \{\ (\varx:\expr)^*\ \}\\
&& \mid&
  \expr \; \kwcat \; \expr & \expr_1~\expr_2\\
&& \mid&
  | \expr | & \mbox{the length of $\expr$}\\
&&\mid&
  \cnstr \kwrl \expr^* \kwrr & \cnstr(\expr^*)\\
&& \mid&
  \varx \kwrl \expr^* \kwrr & \varx~\expr^*\\
&& \mid&
    \wasmc \\

\text{Condition} & \cond &::=&
\multicolumn{2}{l}{
    \kwnot \; \cond \mid
    \cond \; \binop \; \cond \mid
    \expr \; \binop \; \expr \mid
    \kwiscaseof \; \expr \; \varx \mid
    \wasmc
}
\\

\text{Path} & \qual &::=&
\multicolumn{2}{l}{
    \expr \mid
    \expr \bcode{:} \expr \mid
    \bcode{.} x
}
\\
%\text{Wasm} & \wasmcset &\ni& \wasmc \\
  % Wasm-specific condition
  % (* Conditions used in assertions *)
  % | TopLabelC                        (* "a label is now on the top of the stack" *)
  % | TopFrameC                       (* "a frame is now on the top of the stack" *)
  % | TopValueC of expr option (* "a value (of type expr)? is now on the top of the stack" *)
  % | TopValuesC of expr           (* "at least expr number of values on the top of the stack" *)
  \end{array}
\]
\vspace*{-1em}
\caption{Syntax of \al and its prose notation}\label{fig:al-syntax}
\end{figure}

Fig.~\ref{fig:al-syntax} presents the core syntax of \al.
The metavariables $\varx$ ranges over variables,
$\num$ ranges over numbers, and
$\cnstr$ ranges over constructors.
An \al program $\prog$ is a sequence of algorithms $\alg^*$,
which denotes the Wasm semantics in algorithmic style.
An algorithm $\alg$ consists of a name $\varx$, parameters $\expr^*$,
and body instructions $\inst^*$. An algorithm which denotes the semantics of a
certain Wasm instruction or an auxilirary helper function used to describe the language semantics.
An instruction $\inst$ denotes a prose statement in the Wasm specification,
and an expression $\expr$ denotes the expressions that can be evaluated to a value.
A condition $\cond$ denotes an expression that is evaluatd to a boolean value, which can
be used within if instructions or assert instructions.
The figure also shows prose rendering of instructions and expressions.
The Wasm specification often uses specific phrases like ``the current frame'' and
``a label is now on the top of the stack.''
We abstract such Wasm-specific expressions and conditions as $\wasmc$ for brevity.

For example, the semantics of \inblue{\ensuremath{\mathsf{ref.is\_null}}}
in Fig.~\ref{fig:dsl} corresponds to the following in \al:
\[
\small
\!\!
\begin{array}{l}
\kwalg \; \mathit{REF.IS\_NULL} \kwrl \kwrr \; \kwcl \\
\quad \kwassert \; \wasmc \\
\quad \kwpop \; \mathit{val}\\
\quad \kwif \; (\kwiscaseof \; \mathit{val} \; \mathit{REF.NULL})\;
\; (\kwpush \; \ssf{CONST}(\ssf{I32}(), 1)) \; (\kwpush \; \ssf{CONST}(\ssf{I32}(), 0))\\
\kwcr
\end{array}
\]

Due to the nature of AL as a language designed for describing the pseudocode,
the semantics of $\al$ can be defined straightforwardly.
In our companiaon report~\cite{il-tr}, we formally defined the semantics of $\al$ as the state transition system.
Here, we provide a brief overview.
The interpretation of program starts by calling one of its algorithms,
which sequentially executes its body instructions.
Executing one $\al$ instruction alters the program state,
and the final state after executing the last instruction of an algorithm
is the final result of the program execution.

\subsection{\dl to \al Translation}\label{sec:dl2al}
Now, we describe how to translate a Wasm semantics $\delta^*$ in \dl into an \al program $\prog$.
First, the definitions in $\delta^*$ are grouped to represent algorithms; among $\delta^*$,
the reduction rules $\gamma^*$ are grouped according to their target Wasm instructions, and
the helper functions $\lambda^*$ are grouped according to their names.
Each group is translated into a single algorithm in two phases:
1) preprocess the group's definitions to satisfy some \textit{preconditions}, and
2) generate an \al algorithm from the preprocessed \dl definitions.
Here, we will omit the translation of helper functions and
focus on explaining the translation of reduction rules as they share fundamental similarities
with few differences.
The detailed description of the translation is availabe at the companion report~\cite{il-tr}.

\subsubsection{\dl to \dl Preprocessing}
Preprocessing consists of two steps: for each group of reduction rules
1) preprocess the left-hand sides of the reduction rules
in the group to make them the same, and
2) preprocess the premises of the definitions in the group
so that every variable is bound exactly once before its uses.

\medskip
The first preprocessing step is to make the left-hand sides of
the reduction rules in each group identical.
Most Wasm definitions satisfy this precondition, but some do not, such as the following
DSL reduction rules for the \inblue{\ensuremath{\mathsf{br}}} instruction:

{
\begin{verbatim}
     rule Step_pure/br-zero:
        (LABEL_ n `{instr'*} val'* val^n (BR 0) instr*)  ~>  val^n instr'*

     rule Step_pure/br-succ:
        (LABEL_ n `{instr'*} val* (BR $(l+1)) instr*)  ~>  val* (BR l)
\end{verbatim}
}

\noindent
which corresponds to the following in \dl:
\[
\begin{array}{l@{~}l@{~}c@{~}l}
\rho_1 = &(\bot, (\ssf{LABEL\_} \; n \; i'^*\; v'^*\; v^n\; (\ssf{BR\; 0})\; i^*)) &\leadsto&
 (\bot,\; v^n\; i^*)\ \mbox{---}\ []\\
\rho_2 = &(\bot, (\ssf{LABEL\_} \; n \; i'^*\; v^*\; (\ssf{BR\; (}l\ssf{+1)})\; i^*)) &\leadsto&
 (\bot,\; v^*\; (\ssf{BR}\; l))\ \mbox{---}\ []
\end{array}
\]

For each group of reduction rules $\rho^* =
(\eta_1, \eta' _1) \leadsto \gamma_1\ \mbox{---}\ \pi_1^*\
\cdots\
(\eta_n, \eta' _n) \leadsto \gamma_n\ \mbox{---}\ \pi_n^*$,
the unification algorithm \unify takes a list of expressions to unify,
$\eta'_1\ \cdots\ \eta'_n$, and returns a pair of
a unified expression $\eta$ and a list of additional premises $\pi'^*_1 \cdots \pi'^*_n$.
For the \inblue{\ensuremath{\mathsf{br}}} instruction, for example,
\unify takes two expressions
``{$(\ssf{LABEL\_} \; n \; i'^*\; v'^*\; v^n\; (\ssf{BR\; 0})\; i^*)$}'' and
``{$(\ssf{LABEL\_} \; n \; i'^*\; v^*\; (\ssf{BR\; (}l\ssf{+1)})\; i^*$)}.''

For a list of expressions to unify $\eta_1\ \cdots\ \eta_n$,
\unify 1) generates a unified expression $\eta$ possibly containing some fresh variables
and then 2) generates premises $\pi_k^*\ (1\le k \le n)$ using the fresh variables to make
each $\eta_k$ be the same as $\eta$ with $\pi_k^*$.
More specifically, the unified expression $\eta$ is the most common expression of
the expressions to unify $\eta_1\ \cdots\ \eta_n$,
replacing the different components with fresh variables.
For example, the unified expression for the reduction rules for \inblue{\ensuremath{\mathsf{br}}} is
``{$(\ssf{LABEL\_} \; n \; i'^*\; t\; (\ssf{BR}\; t')\; i^*$)}.''
Here, two new variables $t$ and $t'$ are introduced.
After generating the unified expression, \unify generates premises $\pi_k^*$
for each $\eta_k$ so that $\eta_k$ is an instance of the unified expression satisfying $\pi_k^*$.
For example, to make the unified expression for \inblue{\ensuremath{\mathsf{br}}}
same as the left-hand sides of $\rho_1$ and $\rho_2$, \unify infers the conditions
$[t = v'^*\; v^n,\ t' = 0]$ and $[t = v^*,\ t' = l\ssf{+1}]$, respectively.
Finally, the unification result of the rules for the \inblue{\ensuremath{\mathsf{br}}} instruction is as follows:
\[
\begin{array}{l@{~}l@{~}c@{~}l}
\rho_1 = &(\bot, (\ssf{LABEL\_} \; n) \; \eta'^*\; t\; (\ssf{BR}\; t')\; \eta^*) &\leadsto&
 (\bot, v^n\; \eta^*)\ \mbox{---}\ [t = v'^*\; v^n,\ t' = 0]\\
\rho_2 = &(\bot, (\ssf{LABEL\_} \; n) \; \eta'^*\; t\; (\ssf{BR}\; t')\; \eta^*) &\leadsto&
 (\bot, v^*\; (\ssf{BR}\; l))\ \mbox{---}\ [t = v^*,\ t' = l\ssf{+1}]
\end{array}
\]
Thanks to the fresh variables $t$ and $t'$, both rules $\rho_1$ and $\rho_2$ now have
same left-hand sides. The definitions of $t$ and $t'$ are available as premises:
$[t = v'^*\; v^n,\ t' = 0]$ for $\rho_1$ and
$[t = v^*,\ t' = l\ssf{+1}]$ for $\rho_2$.


\medskip
The second preprocessing step is to change the premises of each group's definitions
so that each variable is bound exactly once before it is used.
This step is required because the order of the premises can be
shuffled in arbitrary order in declarative reduction rules.
In addition, an equality expression $\eta = \eta'$ in a premise $\pi$ can be ambiguous,
because it can refer to an equality check condition or a variable binding in either direction.
This second preprocessing step replaces every equality expression denoting a variable binding with $\eta \leftarrow \eta'$.
Thus, this step identifies each variable's binding occurrence,
keeps track of the variables that each premise binds,
and reorders premises so that preceding premises bind all free variables in each premise.

For example, consider the first rule for the \inblue{\ensuremath{\mathsf{br}}} instruction again:
\[
\begin{array}{l@{~}l@{~}c@{~}l}
\rho_1 = &(\bot, (\ssf{LABEL\_} \; n \; i'^*\; t\; (\ssf{BR}\; t')\; i^*)) &\leadsto&
 (\bot, v^n\; \eta^*)\ \mbox{---}\ [t = v'^*\; v^n,\ t' = 0]\\
\end{array}
\]
The second premise $t' = 0$ is an equality check codition,
while the first rule is a binding of fresh variables $v$ and $v'$.
Therefore, this step changes the rule as follows:
\[
\begin{array}{l@{~}l@{~}c@{~}l}
\rho_1 = &(\bot, (\ssf{LABEL\_} \; n \; i'^*\; t\; (\ssf{BR}\; t')\; i^*)) &\leadsto&
 (\bot, v^n\; \eta^*)\ \mbox{---}\ [v'^*\; v^n \leftarrow t,\ t' = 0]\\
\end{array}
\]
which clearly indicates that fresh variables $v$ and $v'$ are newly introduced in this premise.
Once the variable binding occurrences in premises are identified,
reordering the premises is a simple def-use dataflow analysis.

The main problem is now reduced to identifying each variable's binding occurrence in premises,
which turns out to be a \textit{NP-hard} problem.
This is especially because a single premise can bind more than two variables at once,
as in $v'^*\; v^n \leftarrow t$.
We prove it by reduction from a known NP-hard problem, the \textit{exact cover} problem~\cite{exactcover}:
\begin{quote}
The exact cover problem aims at deciding whether it is possible to select some subsets within a given collection of subsets in such a way that each element of a given set belongs to exactly one selected subset. This problem is NP-complete~\cite{karp72}.
\end{quote}
Here, we provide the formal definition of the exact cover problem:
\begin{definition}[Exact Cover Problem]\label{def:exactcover}
An instance of the \textit{Exact Cover Problem} (EC) is defined by a tuple $(X, S)$
such that $X$ is a set of elements and $S \subseteq \mathcal{P}(X)$ is a collection of subsets of $X$.
EC aims at deciding if there exists a subcollection $P \subseteq S$ which is a partition of $X$,
that is, $\forall a \in X.\ |\{ u \in P \mid a \in u \}| = 1$.
\end{definition}

\noindent
For example, consider $X = \{a, b, c, d, e\}$ and $S = \{\{a,b\}, \{b,c\}, \{c,d,e\}\}$.
Because $\{\{a,b\}, \{c,d,e\}\}$, one of the subcollections of $S$, is a partition of $X$,
the answer is yes.
On the other hand, for $S' = \{\{a,b\}, \{b,c\}, \{c,d\}, \{d,e\}\}$,
since no subcollection of $S'$ is a partition of $X$\footnote{It can be easily
verified by the parity. The union of any pair-wise disjoint subset of $S'$ would have an even number
of elements, but the whole set $X$ has 5 elements.}, the answer is no.

Because EC is one of Karp's 21 NP-complete problems~\cite{karp72},
we can prove that the problem of identifying each variable's binding occurrence in premises is NP-hard,
if we can reduce EC into the problem in polynomial time.

\begin{theorem}\label{thm:np-hard}
The problem of identifying each variable's binding occurrence in premises is NP-hard.
\end{theorem}
\begin{proof}
Assume that we are given EC with a set $X$ and
a collection of its subsets $S \subseteq \mathcal{P}(X)$. Let $n$ be the size of $S$.
Let $S_i = \{x_{i1}, x_{i2}, ..., x_{ij}\}$ be the $i$-th subset of $S$.
Now, consider a reduction rule $\gamma \leadsto \gamma'\ \mbox{---}\ \pi^*$
where $\gamma = (\bot,\ v_n, v_{n-1}, \cdots, v_1)$,
$\gamma' = (\bot,\ [])$, and
the $i$-th premise of $\pi^*$ be $v_i = (x_{i1}, x_{i2}, ..., x_{ij})$.
Note that this reduction rule can be constructed in linear time.
The claim is that when we identify each variable's binding occurrence in $\pi^*$,
this gives a solution to EC.
Note that every variable must be bound exactly once.
Thus, if we collect all premises that bind new variables,
then the subsets $S_i$ corresponding to such premises
would form a subcollection of $S$, which is a partition of $X$.
\end{proof}

\textbf{Example.}
Let's consider the example of $X = \{a, b, c, d, e\}$ and $S = \{\{a,b\}, \{b,c\}, \{c,d,e\}\}$ again.
Following the proof above, EC for $X$ and $S$ is reduced to the problem of
identifying each variable's binding occurrence in the premises of the
following reduction rule:
\[
\begin{array}{l@{~}l@{~}c@{~}l}
\rho = & (\bot, [v_3, v_2, v_1]) &\leadsto& (\bot, [])\ \mbox{---}\
[v_1 = (a,b), v_2 = (b,c), v_3 = (c, d, e)]
\end{array}
 \]
If we successfully solve the problem, then the result should look like the following:
\[
\begin{array}{l@{~}l@{~}c@{~}l}
\rho' = & (\bot, [v_3, v_2, v_1]) &\leadsto& (\bot, [])\ \mbox{---}\
[(a,b) \leftarrow v_1, (c, d, e) \leftarrow v_3, v_2 = (b,c)]
\end{array}
\]
From the result, we can reconstruct the partition of the set $X$
by collecting the binding premises, $\{\{a, b\}, \{c, d, e\}\}$, giving the answer to EC.

\begin{algorithm}[t]
\DontPrintSemicolon
\KwIn{A list of premises $\pi^*$ and a set of bound variables $X$}
\KwOut{A list of preprocessed premises $\pi'^*$}
$S \gets \emptyset$\;
\For{$i \gets 1$ \textbf{to} $|\pi^*|$}{
    \lIf{$\pi_i = (\eta = \eta')$}
    {$S_i \gets \{\{i\},\ \{i\}\cup\mathit{free}(\eta),\ \{i\}\cup\mathit{free}(\eta')\}$}
    \lElse{$S_i \gets \{\{i\}\}$}
$S \gets S \cup S_i$\;
}
$P \gets \mathit{Knuth}(S \cup X)$\;
\For{$p \in P$}{
    $\pi'^* \gets \mathit{replaceVariableBinding}(\pi^*, p)$\;
    \lIf{reordering $\pi'^*$ succeeds}
    {\Return{reordered $\pi'^*$}}
}
where $\mathit{free}(\eta)$ returns free variables in $\eta$ and
$\mathit{Knuth}(S)$ returns partitions of $S$
\caption{Preprocess Premises}
\label{algo:preminfer}
\end{algorithm}

\medskip
Thus, no polynomial-time algorithm can solve the problem.
If the numbers of premises are small, a simple brute-force algorithm might be a solution.
However, the reduction rule for module instantiation, for example,
has more than 10 premises and variables, so more efficient method is required.
Another solution is to use an SMT solver like Z3~\cite{z3},
since this problem can be treated as a constraint solving problem.
However, the performance overhead of Z3 does not apply to this problem either.

As a practical solution to the NP-hard problem,
we reduce the problem into EC\footnote{Note that the direction is opposite with
the proof} and then adopt the Knuth algorithm~\cite{knuth2000dancing},
a well-known and effective algorithm for solving EC.
The high-level idea is to encode the premises as a collection of sets,
where a solution to EC of the encoded collection corresponds to the solution to our problem.

Algorithm~\ref{algo:preminfer} describes the process.
It takes two inputs, a list of premises $\pi^*$ in reduction rules
and a set of already bound variables $X$, and returns a list of new premises $\pi'^*$.
First, it encodes premises as a collection of subsets of $X$,
where $X = \{1, \cdots, |\pi^*|\} \cup \mathit{free}(\pi^*)$
is a set of the numbers from $1$ to the size of premises
and all free variables in them, on which the Knuth algorithm performs.
For each $\pi_i$, if it is an equality expression $\eta = \eta'$,
it is encoded as three subsets: $\{\{i\},\ \{i\}\cup\mathit{free}(\eta),\ \{i\}\cup\mathit{free}(\eta')\}$,
which denotes its possible interpretations.
The first set $\{i\}$ denotes when the $i$-th premise does not bind any variables,
meaning $\pi_i$ is an equality check condition.
The second set $\{i\} \cup \mathit{free}(\eta)$ denotes when $\pi_i$ binds all the variables in $\eta$ and
the third set $\{i\} \cup \mathit{free}(\eta')$ denotes when $\pi_i$ binds all the variables in $\eta'$.
For example, if the first premise is $x = y$, it is encoded as three subsets: $\{1\}$, $\{1, x\}$, and $\{1, y\}$.
If $\pi_i$ is not an equality expression, it is encoded as only one subset, $\{i\}$,
meaning $\pi_i$ does not bind any variables but checks some non-equality condition.
Then, the Knuth algorithm takes the collection $S$ containing all the encoded subsets and
the set of bound variables $X$ and returns their partitions $P$.
Note that the Knuth algorithm may not return a unique partion.
In addition, due to the definition of partition, for any partion $p$,
there should be exactly one subset that contains a number $i$ for $1 \le i \le n$.
Thanks to the design of encoding, a subset is either a singleton set $\{i\}$ or
a set with an index and some variables $\{i, x_1, x_2, \cdots\}$.

\begin{algorithm}[t]
\DontPrintSemicolon
\KwIn{A list of premises $\pi^*$ and a partition $p$}
\KwOut{A list of premises $\pi'^*$ with explicit variable binding }
\SetKwProg{Fn}{Function}{}{}
\Fn{$\mathit{replaceVariableBinding}(\pi^*, p)$} {
\For{$i \gets 1$ \textbf{to} $|\pi^*|$}{
\If{$\pi_i = (\eta = \eta')$}
    {\lIf{$\{i\} \in p$}{$\pi'_i \leftarrow (\eta = \eta')$}
     \lElseIf{$\{i\}\cup\mathit{free}(\eta) \in p$}{$\pi'_i \leftarrow (\eta \leftarrow \eta')$}
     \lElseIf{$\{i\}\cup\mathit{free}(\eta') \in p$}{$\pi'_i \leftarrow (\eta' \leftarrow \eta)$}
    }
\lElse{$\pi'_i \leftarrow \pi_i$}
}
\Return{$\pi'^*$}
}
\caption{Replace Variable Binding}
\label{algo:binding}
\end{algorithm}

Using a partition $p$, it replaces every equality expression denoting a variable binding
with an explicit variable binding expression via $\mathit{replaceVariableBinding}$ in Algorithm~\ref{algo:binding}.
For example, if a singleton set $\{2\}$ is in the input partition $p$,
then the second premise $\pi_2$ is a condition and therefore remains the same.
If a set $\{3, y\}$ is in $p$ and the third premise is $x = y$,
then the third premise is replaced with $y \leftarrow x$.
For preprocessed premises $\pi'^*$, Algorithm~\ref{algo:preminfer} tries to reorder them
so that all variables are bound before their uses.
If reordering $\pi'^*$ succeeds, the algorithm returns the reordered $\pi'^*$.
Reordering premises may fail, if they contain cyclic bindings like $x \leftarrow f(y)$ and $y \leftarrow g(x)$.
If that happens, the algorithm tries with the next partition.

\subsubsection{\al Generation}
After preprocessing, \dl definitions satisfy two preconditions:
1) the left-hand sides of the reduction rules in each group
are identical and 2) every variable in premises is bound exactly once before its uses.
For each group of reduction rules $\rho^* =
(\eta_1, \eta' _1)  \leadsto \gamma_1\ \mbox{---}\ \pi_1^*\
\cdots\
(\eta_n, \eta' _n) \leadsto \gamma_n\ \mbox{---}\ \pi_n^*$,
the translation algorithm \dltoil generates an \al algorithm
by translating the left-hand-side $(\eta_1, \eta' _1)$ first,
and then translating the premise $\pi_i$ and right-hand-side $\gamma_i$
for each rule $\rho_i (1\le i\le n)$.
For the rules for \inblue{\ensuremath{\mathsf{ref.is\_null}}} in \dl, for example:
\[
\begin{array}{l@{~}l@{~}c@{~}l}
\rho_1 = &(\bot, [\upsilon, \ssf{REF.IS\_NULL}]) &\leadsto& (\bot, [\ssf{CONST\ I32\ 1}])\
\mbox{---}\ [\upsilon = \ssf{REF.NULL}\ t]\\
\rho_2 = &(\bot, [\upsilon, \ssf{REF.IS\_NULL}])&\leadsto&(\bot, [\ssf{CONST\ I32\ 0}])\
\mbox{---}\ [\ssf{\small otherwise}]
\end{array}
\]
the identical left-hand-side $(\bot, [\upsilon, \ssf{REF.IS\_NULL}])$
is translated to $[\kwassert \; \wasmc, \kwpop \; \mathit{val}]$.
For the first rule $\rho_1$, the premise ``$\upsilon = \ssf{REF.NULL}\ t$'' and
the right-hand-side ``$(\bot, [\ssf{CONST\ I32\ 1}])$'' are translated to
``$\kwif \; (\kwiscaseof \; \mathit{val} \; \mathit{REF.NULL})$'' and
``$\kwpush \; \ssf{CONST}(\ssf{I32}(), 1)$,'' respectively.

Roughly speaking, translating a left-hand-side of a reduction rule corresponds to
generating the beginning of an algorithm, which pops the values from the stack
to use as the inputs for the target instruction and binds new variables
that contain the information about the inputs.
Translation of premises corresponds to the middle of an algorithm,
and generates either variable bindings $\kwlet \; \expr \; \expr$
or if instructions with conditions $\cond$.
Note that a binding expression $\eta \leftarrow \eta'$ may introduce side conditions.
For example, a binding $[x, y] \leftarrow \mathit{arr}$ introduces a side condition
that the size of the array $\mathit{arr}$ is two.
\dltoil generates such side conditions based on binding patterns.
Finally, translation of a right-hand-side of a reduction rule corresponds to
generation of the end of an algorithm, which pushes the result value onto the stack top
or execute other Wasm instructions.
Due to space limitation, we refer the interested readers to a companion report~\cite{il-tr}.
